---
title: "Avocados Price Report"
author: "Malay Raj"
date: "2023-05-20"
output:
  word_document: default
  html_document: default
---

## AVOCADOS PRICE ANALYSIS

: Historical data about avocado prices and sales in grocery stores throughout the United States. Avocado Prices (made available by Justin Kiggins under a CC0 license).

## Loading Required Packages.

```{r}
library(tidyverse)
library(lubridate)
library(ggplot2)
library(plyr)
library(skimr)
```

## Displaying my current working directory.

```{r}
getwd()
```

## Collecting the data.

```{r}
Avocados = read.csv("C:/Avocado Report/avocado.csv")
```


## Displaying the first six rows of the dataset.

```{r}
head(Avocados)
```

## Checking the number of rows and dimension of the dataset.

```{r}
nrow(Avocados)
dim(Avocados) 
```


## Renaming few columns.

```{r}
Avocados=plyr::rename(Avocados, replace=c( "X4046" = "PLU4046",
                                           "X4225" = "PLU4225",
                                           "X4770" = "PLU4770"))
```


## Viewing the new column names.

```{r}
colnames(Avocados)
```


## Inspecting the dataframe and look for inconguencies.

```{r}
str(Avocados)
```


# Convert Date and year to date and numeric datatypes so that they can stack correctly.

```{r}
Avocados$Date = as.Date(Avocados$Date)
Avocados$year = as.numeric(Avocados$year)
```

## Inspecting the dataframe after changing the datatypes.

```{r}
str(Avocados)
```


## Removing the 'X' column because we don't need it in our analysis.

```{r}
Avocados = subset(Avocados, select = -c(X))
```


## Viewing all the column left after removing the column 'X'.

```{r}
colnames(Avocados)
```


## Checking for null values in our dataset.

```{r}
sum(is.null(Avocados))
```

## Checking for any duplicate values in our dataset.

```{r}
anyDuplicated(Avocados)
```


## Seeing how many observations fall under each type.

```{r}
table(Avocados$type)
```


## CONDUCTNG DESCRIPTIVE ANALYSIS:

Descriptive analysis on few columns.

```{r}
summary(Avocados$AveragePrice)
```


```{r}
summary(Avocados$Total.Volume)
```


```{r}
summary(Avocados$PLU4046)
```


```{r}
summary(Avocados$PLU4225)
```


```{r}
summary(Avocados$PLU4770)
```


```{r}
summary(Avocados$Total.Bags)
```


```{r}
summary(Avocados$Total.Bags)
```


```{r}
summary(Avocados$Small.Bags)
```


```{r}
summary(Avocados$Large.Bags)
```


```{r}
summary(Avocados$XLarge.Bags)
```


Comparing AveragePrice and different types.

```{r}
aggregate(Avocados$AveragePrice ~ Avocados$type, FUN = mean)
aggregate(Avocados$AveragePrice ~ Avocados$type, FUN = median)
aggregate(Avocados$AveragePrice ~ Avocados$type, FUN = max)
aggregate(Avocados$AveragePrice ~ Avocados$type, FUN = min)
```


## SUMMARY STATISTICS.

```{r}
skimr::skim_without_charts(Avocados)
```


## DATA VISUALIZATION 

Comparing the AveragePrices of different types.

```{r}
ggplot(Avocados, aes(x = type,y = AveragePrice, color=type)) + 
  geom_boxplot()+
  scale_y_continuous(labels = scales::dollar,limits = c(0,2.5))+
  labs(title="US Total Market: Average Price of one Avocado", x="Date",y="Average Price", color="Type") + 
  theme_minimal()
```


Comparing the AveragePrice of different types of Avocados using histogram.

```{r}
ggplot(Avocados, aes(x = AveragePrice, fill = type)) + 
  geom_histogram(binwidth = 0.20, alpha=0.5)+
  labs(title="Total US Market: Average Price of one Avocado", x="Average Price") + 
  theme_minimal()
```

